---
title: "Final Project: Economic Freedom of the World"
author: "Group1"
date: "2/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Install Packages, message=FALSE, warning=FALSE}
install.packages("caret",repos = "http://cran.us.r-project.org")
pacman::p_load(pacman,janitor,readr,gmodels,tidyr,dplyr,ggthemes,gapminder,plotly,cowplot,tmap,ggplot2,RColorBrewer,gganimate,gifski,sf,factoextra,e1071,cluster,kernlab,corrplot,corrgram,hrbrthemes,viridis,glmnet,rpart,rpart.plot,partykit,maps,tmap,ggmap,htmltools,animation) 

```
## Introduction
For our final project we will be reviewing Economic Freedom across the worls. With data ranging from 1970-2016.

Topics Covered:
  * Exploratory Analysis (Summaries and Filtered Data Frames)
  * Exploratory Plots
  * Linear Regression
  * Variable Selection: Lasso Method
  * Clustering

Step 1: We start by reading in and cleaning the data. In total we have 3726 observations 

```{r Read In Data, message=FALSE, warning=FALSE}
freedom <- read_csv("C:\\Users\\mjone\\OneDrive\\Desktop\\MBA 8045\\Data\\efw_cc.csv") # Update to your file location
freedom <- freedom %>%
  clean_names()

```


Now that we have access to all of our data we will take a look at it to see which variables we have access to
```{r Exploratory Analysis}
str(freedom)#3726 Observations
summary(freedom)
```
# Variable Description:
### Area 1: 
  * Size of Government—As government spending,
taxation, and the size of government-controlled enterprises
increase, government decision-making is substituted for
individual choice and economic freedom is reduced.

### Area 2: 
  * Legal System and Property Rights—Protection of
persons and their rightfully acquired property is a central
element of both economic freedom and civil society.
Indeed, it is the most important function of government.

### Area 3: 
  * Sound Money—Inflation erodes the value of rightfully earned wages and savings. Sound money is thus essential to protect property rights. When inflation is not only high
but also volatile, it becomes difficult for individuals to plan
for the future and thus use economic freedom effectively.

### Area 4: 
  * Freedom to Trade Internationally—Freedom to
exchange—in its broadest sense, buying, selling, making
contracts, and so on—is essential to economic freedom,
which is reduced when freedom to exchange does not
include businesses and individuals in other nations.

### Area 5: 
  * Regulation—Governments not only use a number
of tools to limit the right to exchange internationally, they
may also develop onerous regulations that limit the right to
exchange, gain credit, hire or work for whom you wish, or
freely operate your business. 



### Creating Data Frame for 2016
Next we will create a data frame for 2016 our most recent year. From there we will do some plotting to get a better view of what has occured over time. 
```{r Creating 2016 Data Frame}
freedom2016 <- filter(freedom, year == "2016") #Filtering to 2016 most recent year in data
freedom2016 <- freedom2016[order(freedom2016$rank),]#Sorting data frame from lowest to highest
head(freedom2016)
summary(freedom2016)

#options(repr.plot.width=8, repr.plot.height=10)
```

```{r Highest/Lowest Performers}


```


### Visualize our Data

The below chart has each country in our data set and their score over time 1-10. We can see from our plots that some countries were not established in certain periods. This means we will have to deal with NA values in our analysis. Keep this in mind. 


```{r All Country Scatter Plot, fig.height=7, fig.width=5}

#Plot Below Covers All Countries and Year to get a view of changes
freedom %>% filter(!is.na(economic_freedom) | !is.na(year)) %>% group_by(year) %>%
  ggplot(aes(year,economic_freedom, fill = economic_freedom, group = 1))+
  geom_line(aes(color = economic_freedom))+
  scale_color_viridis(option = "plasma",direction = -1, discrete=FALSE) +
  facet_wrap(~countries)+theme_bw()+
  theme(legend.position = "none", axis.text.x = element_blank(), 
        strip.text.x = element_text(size = 6))+
  xlab(" ") + ylab("")+ ggtitle("ECONOMIC FREEDOM OF THE WORLD 1970-2016")

```


### Scatter Plots

The next three plots show the relation of x (Trade, Sound Money, and Regulations) to economic freedom. The size of the bubble represents the rank. Larger bubbles indicates a higher rank. 

##### Please note a separate file will be send for the video

```{r Plotly Scatter Plot, message=FALSE, warning=FALSE}
#Yearly Scatter Plot for US, Potentially other countries
p <- ggplot(freedom, aes(x4_trade, economic_freedom, color = countries)) +
  geom_point(aes(size = -rank, frame = year, ids = countries)) +
  scale_x_log10()
p <- ggplotly(p)
p


p4 <- ggplot(freedom, aes(x3_sound_money, economic_freedom, color = countries)) +
  geom_point(aes(size = -rank, frame = year, ids = countries)) +
  scale_x_log10()
p4 <- ggplotly(p4)
p4

p5 <- ggplot(freedom, aes(x5c_business_reg, economic_freedom, color = countries)) +
  geom_point(aes(size = -rank, frame = year, ids = countries)) +
  scale_x_log10()
p5 <- ggplotly(p5)
p5
```



## Selected Countries Plot
After reviewing our initial line graphs we selected a few countries to show 6 countries trends. Our x-axis is time and y-axis is Economic Freedom

```{r plot for selected countries, message=FALSE, warning=FALSE}
# Keep only 3 names
country <- filter(freedom,countries %in% c("United States", "China", "Venezuela", "United Kingdom", "Ghana","Zimbabwe"))
 
# Plot
country %>%
  ggplot( aes(x=year, y=economic_freedom, group=countries, color=countries)) +
    geom_line() +
    geom_point() +
    scale_color_viridis(discrete = TRUE) +
    ggtitle("Countries Performance Over Time") +
    theme_ipsum() +
    ylab("Economic Freedom") +
    transition_reveal(year)

# Save at gif:
anim_save("Economic Freedom by Year.gif")

```


Economic Freedom Bubble Plot
Here we can see the distribution of our 

##WORK IN PROGRESS NEED TO UPDATE
```{r Static Bubble Plot}
#static 2016 Scatter Plot
a1 <- ggplotly(ggplot(freedom2016, aes(quartile,economic_freedom , size = -rank)) + 
                 geom_jitter(aes(color=countries, alpha=0.5)) +
                 theme_bw()+ theme(legend.position= "none")+
                 xlab("Quartile") + 
                 ggtitle("Economic Freedom Index 2016"), tooltip = c("countries"))
a1



```



## 2016 Maps (Rank and Trade)
From both of these visualizations we can see that trade and economic_freedom are visually similar to rank. The differences do indicate the other variables should be considered when trying to increase the rank profile of a country.

``` {r 2016 Maps}
jBrewColors <- brewer.pal(n = 8, name = "BuPu")


p1 <- plot_geo(freedom2016) %>%
  add_trace(z = ~-rank, color = ~-rank,colors = jBrewColors,
            text = ~freedom2016$countries, locations = ~freedom2016$iso_code) %>%
  colorbar(title = 'Countries' , tickprefix = 'Rank') %>%
  layout(title = 'Economic Freedom 2016 by Rank') 

p1

p2 <- plot_geo(freedom2016) %>%
  add_trace(z = ~x5c_business_reg, color = ~x5c_business_reg,colors = jBrewColors,
            text = ~freedom2016$countries, locations = ~freedom2016$iso_code) %>%
  colorbar(title = 'Countries' , tickprefix = 'Trade') %>%
  layout(title = 'Economic Freedom 2016 by Trade') 

p2

p8 <- plot_geo(freedom2016) %>%
  add_trace(z = ~economic_freedom, color = ~-rank,colors = jBrewColors,
            text = ~freedom2016$countries, locations = ~freedom2016$iso_code) %>%
  colorbar(title = 'Countries' , tickprefix = 'Rank') %>%
  layout(title = 'Economic Freedom 2016 by Rank') 

p8

```


## Assigning Clusters For 2016 
```{r 2016 Cluster}
fviz_nbclust(na.omit(freedom2016[4:36]), kmeans, method = "silhouette")
gap_stat <- clusGap(na.omit(freedom2016[4:36]), FUN = kmeans, nstart = 25,
                    K.max = 10, B = 50)
fviz_gap_stat(gap_stat)

n_cluster <- 4

results  <- kmeans(na.omit(freedom2016[4:36]),n_cluster)
results

freedom2016 <- na.omit(freedom2016)
freedom2016$cluster_number <- results$cluster
```

## Assigning Clusters for All Years 
```{r Cluster}
fviz_nbclust(na.omit(freedom[4:36]), kmeans, method = "silhouette")
gap_stat <- clusGap(na.omit(freedom[4:36]), FUN = kmeans, nstart = 25,
                    K.max = 10, B = 50)
fviz_gap_stat(gap_stat)

n_cluster <- 4

results2  <- kmeans(na.omit(freedom[4:36]),n_cluster)
results2

freedom <- na.omit(freedom)
freedom$cluster_number <- results2$cluster
```


# Cluster Plots 2016
Note to Megan correct the color to be specific to each cluster not scale
```{r 2016 Cluster Plot, message=FALSE, warning=FALSE}

fviz_cluster(results, data = na.omit(freedom2016[4:36]))

jBrewColors2 <- brewer.pal(n = 8, name = "Spectral")
p3 <- plot_geo(freedom2016) %>%
  add_trace(z = ~cluster_number, color = ~cluster_number,colors = jBrewColors2,
            text = ~freedom2016$countries, locations = ~freedom2016$iso_code) %>%
  colorbar(title = 'Countries' , tickprefix = 'Clusters') %>%
  layout(title = 'Economic Freedom 2016 by Cluster') 

p3

```
# Cluster Plots All Years
Note to Megan correct the color to be specific to each cluster not scale. Will only include if can get animation to work
```{r Cluster Plot, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

fviz_cluster(results2, data = na.omit(freedom[4:36]))

jBrewColors2 <- brewer.pal(n = 8, name = "Spectral")
p3 <- plot_geo(freedom) %>%
  add_trace(z = ~cluster_number, color = ~cluster_number,colors = jBrewColors2,
            text = ~freedom$countries, locations = ~freedom$iso_code) %>%
  colorbar(title = 'Countries' , tickprefix = 'Clusters') %>%
  layout(title = 'Economic Freedom 2016 by Cluster') 

p3

```


# Animated Map For Cluster
Chunck Above trying to solve the same problem
```{r eval=FALSE, include=FALSE}

head(freedom)

saveGIF({
for(i in seq(freedom[4:36])){
m <- freedom%>% filter(year==i) %>% plot_geo(freedom[4:36]) %>% add_trace(z = freedom$cluster_number, color = freedom$cluster_number, colors = jBrewColors2,
            text = ~freedom$countries, locations = ~freedom$iso_code) %>%
  colorbar(title = 'Countries' , tickprefix = 'Cluster') %>%
  layout(title = 'Economic Freedom 2016 by Cluster')  + xlab("LONGITUDE") + ylab("LATITUDE") + theme(legend.position='top') + scale_color_gradient(low = "red", high = "green") + ggtitle(paste0("YEAR : ",i))

print(m)
    }
},interval = .5,movie.name = 'map.gif',ani.width = 1200, ani.height = 800)



```


# Regression

Now we get into the meat of our analysis. We start be prepping our data for regression. Gathering some test and training data

```{r Regression}
str(freedom)
freedom$high_ef <- ifelse(freedom$economic_freedom > 6, 1, 0) # Converting streams to binary data 
freedom <- na.omit(freedom)
index_training <- sample(dim(freedom)[1], 40, replace = FALSE) # replace = false will prevent code from selecting same observation twice
training_data  <- freedom[index_training, ]
test_data      <- freedom[-index_training, ]

```


Since we dont want to randomly select variables to use we will use the Lasso Method for variable selection

```{r Lasso Method}
typeof(training_data[, 7:36])
X_training <- data.matrix(training_data[, 7:36])
X_test     <- data.matrix(test_data[, 7:36])

model_linear_lasso <- cv.glmnet(X_training, training_data$economic_freedom, family="gaussian", nfolds = 10, type.measure = "mse", alpha = 1)
summary(model_linear_lasso)


model_linear_lasso$lambda

coef(model_linear_lasso, s = 0.4)
coef(model_linear_lasso, s = 500)

as.double(coef(model_linear_lasso, s = 0.4))
as.double(coef(model_linear_lasso, s = 500))

#After Lasso Linear Regression
mlr2 <- lm(economic_freedom ~ x3_sound_money + x4_trade + x5c_business_reg, data = training_data)
anova(mlr2)

```


Now that we have our predicted values for our model we want to check how our work.

```{r Testing Prediction vs. Actuals}


```

Provide insights by cluster and prediction points

```{r Cluster Regression Analysis}




```
```{r ROC Plot}


```


Finally we plot our linear regression with the variables selected from the variable selection
```{r Final Regression Plot}
plot_grid(ggplot(freedom, aes(x3_sound_money, economic_freedom)) + 
            geom_point(color='#50C878', alpha = 0.5)+
            xlab("Sound Money") + ylab("Economic Freedom ")+
            geom_smooth(method = lm)+theme_bw(),
          ggplot(freedom, aes(x4_trade, economic_freedom)) + 
            geom_point(color='#50C878', alpha = 0.5)+
            xlab("Trade") + ylab("")+
            geom_smooth(method = lm)+theme_bw(),
          ggplot(freedom, aes(x5c_business_reg, economic_freedom)) + 
            geom_point(color='#50C878', alpha = 0.5)+
            xlab("Business_Regulations") + ylab("")+
            geom_smooth(method = lm)+theme_bw(),align = 'h')
```


# Conclusion/ Key Takeaways
Note to Group: We can certainly change these as everyone becomes familiar with the data and forms their own takeaways

Dealing with a long time frame created the challenge of dealing with null values. 

Cluster 2 had the key attributes of countries that ranked the highests.

Since the aspects of Economic Freedom have been broken down into measurable variables we were able to build sound models

Our models proved to be significant in predicting economic freedom or rank (update this later). If government officials want to improve the lives of their citizens through economic freedom they should focus on:
  * Having sound money
  * Free Trade
  * Business Regulation

## Thank you for working with us this semester!

Sources:
https://www.kaggle.com/gsutters/economic-freedom

https://www.fraserinstitute.org/sites/default/files/economic-freedom-of-the-world-2018-execsummary.pdf