---
title: "Final Project"
author: "Group1"
date: "2/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Install Packages, message=FALSE, warning=FALSE}
install.packages("caret",repos = "http://cran.us.r-project.org")
pacman::p_load(pacman,janitor,readr,gmodels,tidyr,dplyr,ggthemes,gapminder,plotly,cowplot,tmap,ggplot2,RColorBrewer,gganimate,gifski,sf,factoextra,e1071,cluster,kernlab,corrplot,corrgram,hrbrthemes,viridis,glmnet,rpart,rpart.plot,partykit,maps,tmap,ggmap,htmltools) 

```

For our final project we will be reviewing Economic Freedom across the worls. With data ranging from 1970-2016. We start by reading in and cleaning the data. In total we have 3726 observations 

```{r Read In Data, message=FALSE, warning=FALSE}
freedom <- read_csv("C:\\Users\\mjone\\OneDrive\\Desktop\\MBA 8045\\Data\\efw_cc.csv") # Update to your file location
freedom <- freedom %>%
  clean_names()

```


Now that we have access to all of our data we will take a look at it to see which variables we have access to
```{r Exploratory Analysis}
str(freedom) #3726 Observations
```





Next we will create a data frame for 2016 our most recent year. From there we will do some plotting to get a better view of what has occured over time. 
```{r}
freedom2016 <- filter(freedom, year == "2016") #Filtering to 2016 most recent year in data
freedom2016 <- freedom2016[order(freedom2016$rank),]#Sorting data frame from lowest to highest
head(freedom2016)

#options(repr.plot.width=8, repr.plot.height=10)


#Plot Below Covers All Countries and Year to get a view of changes
freedom %>% filter(!is.na(economic_freedom) | !is.na(year)) %>% group_by(year) %>%
  ggplot(aes(year,economic_freedom, fill = economic_freedom, group = 1))+
  geom_line(aes(color = economic_freedom))+
  scale_color_viridis(option = "plasma",direction = -1, discrete=FALSE) +
  facet_wrap(~countries)+theme_bw()+
  theme(legend.position = "none", axis.text.x = element_blank(), 
        strip.text.x = element_text(size = 6))+
  xlab(" ") + ylab("")+ ggtitle("ECONOMIC FREEDOM OF THE WORLD 1970-2016")

```

```{r}
#static 2016 Plot
a1 <- ggplotly(ggplot(freedom2016, aes(quartile,economic_freedom , size = -rank)) + 
                 geom_jitter(aes(color=countries, alpha=0.5)) +
                 theme_bw()+ theme(legend.position= "none")+
                 xlab("Quartile") + 
                 ggtitle("Economic Freedom Index 2016"), tooltip = c("countries"))
a1

#Yearly Scatter Plot for US, Potentially other countries
p <- ggplot(freedom, aes(x1a_government_consumption, economic_freedom, color = countries)) +
  geom_point(aes(size = -rank, frame = year, ids = countries)) +
  scale_x_log10()
p <- ggplotly(p)
p


```


``` {r}
jBrewColors <- brewer.pal(n = 8, name = "BuPu")


p1 <- plot_geo(freedom2016) %>%
  add_trace(z = ~-rank, color = ~-rank,colors = jBrewColors,
            text = ~freedom2016$countries, locations = ~freedom2016$iso_code) %>%
  colorbar(title = 'Countries' , tickprefix = 'Rank') %>%
  layout(title = 'Economic Freedom 2016 by Rank') 

p1

p2 <- plot_geo(freedom2016) %>%
  add_trace(z = ~x5c_business_reg, color = ~x5c_business_reg,colors = jBrewColors,
            text = ~freedom2016$countries, locations = ~freedom2016$iso_code) %>%
  colorbar(title = 'Countries' , tickprefix = 'Trade') %>%
  layout(title = 'Economic Freedom 2016 by Trade') 

p2


```


#Assigning Clusters Here
```{r}
fviz_nbclust(na.omit(freedom2016[4:36]), kmeans, method = "silhouette")
gap_stat <- clusGap(na.omit(freedom2016[4:36]), FUN = kmeans, nstart = 25,
                    K.max = 10, B = 50)
fviz_gap_stat(gap_stat)

n_cluster <- 4

results  <- kmeans(na.omit(freedom2016[4:36]),n_cluster)
results

freedom2016 <- na.omit(freedom2016)
freedom2016$cluster_number <- results$cluster
```


#Cluster Plots
Note to Megan correct the color to be specific to each cluster not scale
```{r}
fviz_cluster(results, data = freedom2016[4:36])


jBrewColors2 <- brewer.pal(n = 8, name = "Spectral")
p3 <- plot_geo(freedom2016) %>%
  add_trace(z = ~cluster_number, color = ~cluster_number,colors = jBrewColors2,
            text = ~freedom2016$countries, locations = ~freedom2016$iso_code) %>%
  colorbar(title = 'Countries' , tickprefix = 'Clusters') %>%
  layout(title = 'Economic Freedom 2016 by Cluster') 

p3

```


#Regression
Adjust to include Clusters
Now we get into the meat of our analysis. We start be prepping our data for regression. Gathering some test and training data

```{r Regression}
str(freedom)
freedom$high_ef <- ifelse(freedom$economic_freedom > 6, 1, 0) # Converting streams to binary data 
freedom <- na.omit(freedom)
index_training <- sample(dim(freedom)[1], 40, replace = FALSE) # replace = false will prevent code from selecting same observation twice
training_data  <- freedom[index_training, ]
test_data      <- freedom[-index_training, ]

```
Provide insights by cluster

Since we dont want to randomly select variables to use we will use the Lasso Method for variable selection

```{r Lasso Method}
typeof(training_data[, 7:36])
X_training <- data.matrix(training_data[, 7:36])
X_test     <- data.matrix(test_data[, 7:36])

model_linear_lasso <- cv.glmnet(X_training, training_data$economic_freedom, family="gaussian", nfolds = 10, type.measure = "mse", alpha = 1)
summary(model_linear_lasso)


model_linear_lasso$lambda

coef(model_linear_lasso, s = 0.4)
coef(model_linear_lasso, s = 500)

as.double(coef(model_linear_lasso, s = 0.4))
as.double(coef(model_linear_lasso, s = 500))

#After Lasso Linear Regression
mlr2 <- lm(economic_freedom ~ x3_sound_money + x4_trade + x5c_business_reg, data = training_data)
anova(mlr2)
plot(mlr2)
```

Provide insights by cluster and prediction points

Finally we plot our linear regression with the variables selected from the variable selection
```{r Final Regression Plot}
plot_grid(ggplot(freedom, aes(x3_sound_money, economic_freedom)) + 
            geom_point(color='#50C878', alpha = 0.5)+
            xlab("Sound Money") + ylab("Economic Freedom ")+
            geom_smooth(method = lm)+theme_bw(),
          ggplot(freedom, aes(x4_trade, economic_freedom)) + 
            geom_point(color='#50C878', alpha = 0.5)+
            xlab("Trade") + ylab("")+
            geom_smooth(method = lm)+theme_bw(),
          ggplot(freedom, aes(x5c_business_reg, economic_freedom)) + 
            geom_point(color='#50C878', alpha = 0.5)+
            xlab("Business_Regulations") + ylab("")+
            geom_smooth(method = lm)+theme_bw(),align = 'h')
```


Sources:
https://www.kaggle.com/gsutters/economic-freedom

https://www.fraserinstitute.org/sites/default/files/economic-freedom-of-the-world-2018-execsummary.pdf